---
title: "Homework_08"
format: html
---

March 4 2026
For loops and randomization tests 
Gwen Phillips
#############################################################################

Q1. Using a for loop, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a counter variable counter <- 0. Inside the loop, add 1 to counter each time you have a zero in the vector. Finally, use return(counter) for the output.

```{r}
 num_zeroes <- function(numeric_vector) {
  # Set up a counter variable before entering the loop
  counter <- 0
 
 for (i in seq_along(numeric_vector)) {
    if (numeric_vector[i] == 0) {
        # If it is, add 1 to the counter
        counter <- counter + 1}
  }
  return(counter)
 }

my_vector <- sample(0:25, 100, replace= TRUE)
zero_count <- num_zeroes(my_vector)
print(zero_count)

```

Q2. Use subsetting instead of a loop to rewrite the function as a single line of code.

```{r}
q2 <- sum(my_vector == 0)
print (q2)
```

Q3. Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.

```{r}
matrix <- function(rows, cols) {
    i <- 1:rows
    j <- 1:cols
    ### multiply the row and columns
        result_matrix <- outer(i, j, "*")
    
    return(result_matrix)
}

```

Q4. In the next few lectures, you will learn how to do a randomization test on your data. We will complete some of the steps today to practice calling custom functions within a for loop.
a. Simulate a dataset with 3 groups of data, each group drawn from a distribution with a different mean. The final data frame should have 1 column for group and 1 column for the response variable.

```{r}
### Set seed so nums don't change later
set.seed(123)

# Generate data for 3 groups
n <- 10
df <- data.frame(
  Group = rep(c("1", "2", "3"), each = n),
  Value = c(rnorm(n, mean = 7),
            rnorm(n, mean = 13),
            rnorm(n, mean = 18))
)

head(df)
```

b. Write a custom function that 1) reshuffles the response variable, and 2) calculates the mean of each group in the reshuffled data. Store the means in a vector of length 3.

```{r}
shuffle <- function() {
  group1 <- rnorm(10, mean = 7)
  group2 <- rnorm(10, mean = 13)
  group3 <- rnorm(10, mean = 18)
  
  # Return the means of the 3 groups
  return(c(mean(group1), mean(group2), mean(group3)))
}



```

c. Use a for loop to repeat the function in b 100 times. Store the results in a data frame that has 1 column indicating the replicate number and 1 column for each new group mean, for a total of 4 columns.

```{r}
n_sim <- 100

results_df <- data.frame(
  replicate = 1:n_sim,
  mean1 = NA,
  mean2 = NA,
  mean3 = NA
)

for (i in 1:n_sim) {
  # Run function b
  rep_means <- shuffle()
  
  # Store results in the respective columns
  results_df[i, "mean1"] <- rep_means[1]
  results_df[i, "mean2"] <- rep_means[2]
  results_df[i, "mean3"] <- rep_means[3]
}
```


d. Use qplot() to create a histogram of the means for each reshuffled group. Or, if you want a challenge, use ggplot() to overlay all 3 histograms in the same figure. How do the distributions of reshuffled means compare to the original means?

```{r}
library(ggplot2)
library(reshape2)
library(patchwork)

g1<- ggplot(results_df, aes(x = mean1)) +
  geom_histogram(position = "identity", alpha = 0.5, binwidth = 0.5) +
  labs(title = "Shuffled Means", x = "Mean", y = "Frequency")

g2<- ggplot(results_df, aes(x = mean2)) +
  geom_histogram(position = "identity", alpha = 0.5, binwidth = 0.5) +
  labs(title = "Shuffled Means", x = "Mean", y = "Frequency")

g3<- ggplot(results_df, aes(x = mean3)) +
  geom_histogram(position = "identity", alpha = 0.5, binwidth = 0.5) +
  labs(title = "Shuffled Means", x = "Mean", y = "Frequency")

g1
g2
g3

 g1 + g2 + g3

### These are all normal distributions, they are shuffled because no seed was set but the means remained 7, 13 and 18.
```

