---
title: "Homework_09"
format: html
---

### Homework 9
### 18 March 2026
### Gwen Phillips

#############################################################

load in packages

```{r}
library(ggplot2)
library(dplyr)
d<- iris
```

Q1. Create
Scatter plot of (y) vs (x)
Regression line using geom_smooth(method = "lm")

```{r}
head(d)
ggplot(data=d)+
  aes(x=Petal.Length, y= Petal.Width)+
  geom_point() +
    labs(
    title = "Petal Width vs Petal Length",
    x = "Petal Length", y = "Petal Width")+
  geom_smooth() +
  geom_smooth(method = "lm", col= "red")
```

Q2. Create
Boxplot or violin plot
Points overlaid using jitter (geom_jitter())


```{r}
ggplot(iris, aes(x = Species, y = Petal.Length, fill = Species)) +
  geom_violin(alpha = 0.5) +
   labs(title = "Species vs Petal Length", y= "Petal Length")+
  geom_jitter(width = 0.2, alpha = 0.7)
    
```

Q3. Create 
Jittered points
Logistic regression curve using geom_smooth(method = "glm", method.args = list(family = "binomial"))

```{r}
d2<- filter(iris, Species!="setosa")
library(ggplot2)

d3 <- d2 %>%
  mutate(binary = if_else(Species == "virginica", 1, 0))
head(d3)

ggplot(d3, aes(x = Petal.Length, y = binary)) +
  geom_jitter(height = 0.05, width = 0, alpha = 0.5) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"), 
              se = TRUE) +
  labs(title = "Logistic Regression with Jittered Data",
       x = "Petal Length)",
       y = "Species (Versicolor=0, Virginica=1)") 
```

Q4. Create a mosaic plot using two discrete variables

```{r}
d4 <- d3 %>%
  mutate(binarypetallength = if_else(Petal.Length < "5.5", 1, 0))
head(d4)

```