---
title: "Homework 13"
format: html
---

Q1: How does life expectancy change with gdp


Answer: I expected life expectancy to increase per gdp with increased values in developed continents. GDP is a common way to predict f a country has higher life expectancy. There is very large variation between richer and poorer countries but there appears to be a relationship.
```{r}
# Load libraries
library(gapminder)
library(dplyr)
library(ggplot2)

head(gapminder)

### How does life expectancy change with gdp

model <- glm(lifeExp ~ gdpPercap * continent,
             family = Gamma(link = "log"),
             data = gapminder)

summary(model)

ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, color = continent)) +
  geom_point(alpha = 0.4)  +
  scale_x_log10() +
  labs(
    title = "Gamma GLM: Life Expectancy vs GDP per Capita by Continent",
    x = "GDP per Capita (log scale)",
    y = "Life Expectancy"
  ) +
  theme_minimal()

```


Q2: How does population affect gdp

Answer: There does not appear to be a strong relationship between gdp and population. There are some countries that have very high gdp per populations, especially in Asia. This was not a relationship I expected to be strong.
```{r}
library(gapminder)
library(lme4)
library(dplyr)

data <- gapminder %>%
  mutate(
    log_pop = log(pop),
    log_gdpPercap = log(gdpPercap)
  )

model2 <- lmer(
  log_gdpPercap ~ log_pop + year + continent + (1 | country),
  data = data
)

summary(model2)

ggplot(data, aes(x = log_pop, y = gdpPercap, color = continent)) +
  geom_point(alpha = 0.4)  +
  scale_x_log10() +
  labs(
    title = "GDP per capita by continent vs log of population",
    x = "Population (log)",
    y = "GDP per capita"
  ) +
  theme_minimal()
```