---
title: "Homework_04"
format: html
---

### Homework 4
### Gwen Phillips
### Feb 4 2026

###________________________

Q1.
The sample function provides a random number within the constraints provided.  The t() function transposes the matrix or data frame referenced within the parenthesis.  To calculate the sum and mean of certain rows utilize the brackets within the fuction. 

```{r}
n_dims<- sample(3:10, 1)
my_vec<- 1:(n_dims^2)

s<- sample(my_vec, 4)
m<- matrix(s, nrow=2, ncol=2)
print(m)

###transposing the matrix
t(m)
print(m)

### finding matrix sums and means
sum(m[1,])
mean(m[1,])
sum(m[2,])
mean(m[2,])

###eigen function
help(eigen)
e<- eigen(m,symmetric=T)
print(e)
#### These are floats since they have decimals
typeof(e)
### These are considered a list
```

Q2.
Create matrixes, vectors, and lists, then identify them using the typeof() function.  Then isolate certain values within those and creat lists and vectors.

```{r}
### create a matrix
my_matrix<- matrix(runif(16),nrow=4, ncol=4)
my_matrix
### create a T/F vector
v<- sample(1:1000, 100)
my_logical<- v>=500
my_logical
### create a vector of 26 random letters
my_letters<- sample(letters, 26)
my_letters

### create a list 
list<-list(my_matrix[2,2],my_logical[2],my_letters[2])

print(list)
### confirm the list
typeof(list) ###it is a list

### create a vec of the list elements
vec<- c(0.01371653, "TRUE", "l")
### what data type is it
typeof(vec) ### character
```

Q3
Set.seed function allows random numbers to be saved so they don't change in later formulas.  Data frames can be created using the cbind function and the which function can highight specific elements of the data.  The order function reorders that data frame in numeric or alphabetical order depending on the type of vectors.
```{r}
### create 26 values between 0 and 10
set.seed(321)
my_unis<- runif(26, min=0, max=10)
my_unis

### create a vector of 26 letters
set.seed(123)
my_letters<- sample(LETTERS, 26)
my_letters

### create a df
df<- cbind(my_unis, my_letters)
### Enter NAs into 4 random rows
df [sample(nrow(my_unis),4),]<- NA ### didnt work

### intentify which rows have a missing variable
which(!complete.cases(df))
### reorder to aphabetical
df_alphabetic<- df[order(df[,2] )]### numbers in alphabet order but doesn't show letters?
print(df_alphabetic)
### find mean of col 1
mean_df<- mean(df[,1]) ### not working idk
print(mean_df)
```


